---
title: "Tennsesse-Statewide land Use Dashboard"
author: ""
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    social: menu
---

```{r setup, include=FALSE}
library(ggplot2)
library(plotly)
library(plyr)
library(flexdashboard)
library(leaflet)
library(sf)
library(readr)
library(dplyr)
library(ggplot2)
library(plotly)
library(tidyr)
library(readxl)

# create some data
# Load the shapefile
shape_data <- st_read("./TAZ_Shapefile/TN_TAZ_2010.shp")

shape_data <- st_transform(shape_data, 4326)

shape_data <- st_centroid(shape_data)
# Load the CSV file


csv_data <- read_excel("./Forecasts_test/Forecasting_Result_2045_TDM.xlsx")

# Ensure the key columns are of the same data type
csv_data$TAZ_ID <- as.integer(csv_data$TAZ_ID)

# Merge the data
merged_data <- left_join(shape_data, csv_data, by = c("TAZID" = "TAZ_ID"))
```

# Population{.tabset}

## charts{.tabset}

### Graphs

```{r}

aggregated_data <- merged_data %>%
  group_by(REGION) %>%
  summarise(GQPOP = sum(GQPOP, na.rm = TRUE))

GQPOP<-ggplot(aggregated_data, aes(x = REGION, y = GQPOP, fill = REGION)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Group Quarters Population by REGION",
       x = "REGION",
       y = "Group Quarters Population") +
  coord_flip() # For horizontal bars
ggplotly(GQPOP)

```


### Mean Population types by region

```{r}

aggregated_data <- merged_data %>%
  group_by(REGION) %>%
  summarise(TOTPOP = mean(TOTPOP, na.rm = TRUE),
            HHPOP = mean(HHPOP, na.rm = TRUE),
            GQPOP = mean(GQPOP, na.rm = TRUE))
# Reshape data for plotting with ggplot2
long_data <- aggregated_data %>% 
  pivot_longer(cols = c("TOTPOP", "HHPOP", "GQPOP"), names_to = "Population_Type", values_to = "Value")

# Create the bar plot
BAR<-ggplot(long_data, aes(x = REGION, y = Value, fill = Population_Type)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  theme_minimal() + 
  labs(title = "Mean Population Types by REGION", x = "REGION", y = "Mean Population") + 
  scale_fill_brewer(palette = "Set2") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggplotly(BAR)



```



### Region Map

```{r}


# Generate a simple color palette
num_regions <- length(unique(merged_data$REGION))
palette <- colorFactor(palette = "viridis", domain = merged_data$REGION)

leaflet(merged_data) %>%
  addTiles() %>%
  addCircleMarkers(
    fillColor = ~palette(REGION),
    color = NA,  # Set border color to transparent
   fillOpacity = 0.7,  radius = 7,
    label = ~as.character(REGION),
    labelOptions = labelOptions(
      style = list("font-weight" = "normal"),
      textsize = "13px",
      direction = "auto"
    )
  ) %>%
  addLegend("bottomright", pal = palette, values = ~REGION,
            title = "Region",
            opacity = 0.7)




```



### Total Population distribution

```{r}

popPalette <- colorNumeric(palette = "Oranges", domain = merged_data$TOTPOP)

leaflet(merged_data) %>% 
  addTiles() %>% 
  addCircleMarkers(
    fillColor = ~popPalette(TOTPOP),  # Color by total population
    color = NA,  # Set border color to transparent
    fillOpacity = 0.7,  # Fill opacity
    radius = 2.5,
    label = ~paste("Population:", TOTPOP),  # Show population on hover
    labelOptions = labelOptions(
      style = list("font-weight" = "normal"),
      textsize = "13px",
      direction = "auto"
    ),
    group = "2018"
  ) %>%
  addLayersControl(
    overlayGroups = c("2018"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  addLegend(
    position = "bottomright",
    pal = popPalette,
    values = ~TOTPOP,
    title = "Total Population",
    opacity = 0.7
  )


```

## Row

### ALL populations Vs MPO

```{r}

aggregated_data <- merged_data %>%
  group_by(MPO) %>%
  summarise(TOTPOP = sum(TOTPOP, na.rm = TRUE),
            HHPOP = sum(HHPOP, na.rm = TRUE),
            GQPOP = sum(GQPOP, na.rm = TRUE))
# Reshape data for plotting with ggplot2
long_data <- aggregated_data %>% 
  pivot_longer(cols = c("TOTPOP", "HHPOP", "GQPOP"), names_to = "Population_Type", values_to = "Value")

# Create the bar plot
BAR<-ggplot(long_data, aes(x = MPO, y = Value, fill = Population_Type)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  theme_minimal() + 
  labs(title = "Population Types by MPO", x = "MPO", y = "Population") + 
  scale_fill_brewer(palette = "Set2") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggplotly(BAR)

```

### Group Quarters vs MPO

```{r}


# Generate a color palette function
aggregated_data <- merged_data %>%
  group_by(MPO) %>%
  summarise(GQPOP = sum(GQPOP, na.rm = TRUE))

aggregated_data <- aggregated_data %>%
  arrange(desc(GQPOP))

GQPOP_plot <- ggplot(aggregated_data, aes(x = reorder(MPO, GQPOP), y = GQPOP, fill = MPO)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Group Quarters Population by MPO",
       x = "MPO",
       y = "Group Quarters Population") +
  coord_flip() # For horizontal bars

ggplotly(GQPOP_plot)


```

### Population

```{r}

continous<-ggplot(merged_data, aes(x = TOTPOP)) +
  geom_density(alpha = 0.3, linetype = "solid", color = "red", size = 1.5, fill = "red") +  # Added fill and increased size
  geom_histogram(aes(y = ..density..), binwidth = 100, fill = "blue", color = "black", alpha = 0.5) +  # Use density scale for y
  labs(title = "Total Population Distribution", x = "Total Population", y = "Density") +
  theme_minimal()
ggplotly(continous)


```



# HouseHold{.tabset}

## charts{.tabset}

### Graphs

```{r}

# Define a color for each household size group
colors <- c('1' = "#1f77b4", '2' = "#ff7f0e", '3' = "#2ca02c", 
            '4' = "#d62728", '5' = "#9467bd", '6' = "#8c564b")
data <- merged_data %>%
  mutate(HHSIZE_group = case_when(
    HHSIZE >= 0 & HHSIZE <= 1.5 ~ '1',
    HHSIZE > 1.5 & HHSIZE <= 2.5 ~ '2',
    HHSIZE > 2.5 & HHSIZE <= 3.5 ~ '3',
    HHSIZE > 3.5 & HHSIZE <= 4.5 ~ '4', 
    HHSIZE > 4.5 & HHSIZE <= 5.5 ~ '5',
    HHSIZE > 5.5 ~ '6'
  ))

p <- ggplot(data, aes(x = HHSIZE_group, fill = HHSIZE_group)) +
  geom_bar(stat = "count", color = "black") +
  scale_fill_manual(values = colors, 
                    name = "Household Size Group", 
                    labels = c('1' = "0-1.5", '2' = "1.5-2.5", '3' = "2.5-3.5", 
                               '4' = "3.5-4.5", '5' = "4.5-5.5", '6' = "5.5+")) +
  labs(title = "Frequency of Household Size Groups",
       x = "Household Size Group",
       y = "Frequency") +
  theme_minimal()

# Convert to an interactive plot with plotly
ggplotly(p)


```

### AVG HHINC vs MPO

```{r}


aggregated_data <- merged_data %>%
  filter(!is.na(MPO)) %>%
  group_by(MPO) %>%
  summarise(HHINC = mean(HHINC, na.rm = TRUE))

HHI<-ggplot(aggregated_data, aes(x = MPO, y = HHINC, fill = MPO)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "HouseHold Income by MPO",
       x = "MPO",
       y = "HouseHold Income") +
  coord_flip() # For horizontal bars
ggplotly(HHI)


```


## Row

### AVG Household's in an MPO

```{r}

aggregated_data <- merged_data %>%
  filter(!is.na(MPO)) %>%
  group_by(MPO) %>%
  summarise(HH = mean(HH, na.rm = TRUE))

HH<-ggplot(aggregated_data, aes(x = MPO, y = HH, fill = MPO)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "HouseHold by MPO",
       x = "MPO",
       y = "HouseHold") +
  coord_flip() # For horizontal bars
ggplotly(HH)

```

### Region Map

```{r}
library(ggplot2)
library(plotly)

# Assuming 'merged_data' is prepared and contains 'TOTEMP'

# Create the ggplot object
ggplot_map <- ggplot(data = merged_data) +
  geom_sf(aes(color = HHINC), size = 1.5) +  # Use color to represent TOTEMP
  scale_color_distiller(palette = "Spectral", direction = 1) +  # Adjust palette as needed
  labs(title = "Total Employment",
       subtitle = "By Traffic Analysis Zone or MPO",
       color = "Total Employment") +
  theme_minimal() +
  theme(legend.position = "bottom")

# Convert the ggplot object to plotly
interactive_map <- ggplotly(ggplot_map)

# Display the interactive map
interactive_map



```

### Household Income

```{r}

continous<-ggplot(merged_data, aes(x = HHINC)) +
  geom_density(alpha = 0.3, linetype = "solid", color = "red", size = 1.5, fill = "red") +  # Added fill and increased size
  geom_histogram(aes(y = ..density..), binwidth = 100, fill = "blue", color = "black", alpha = 0.5) +  # Use density scale for y
  labs(title = "Household Income Distribution", x = "Household Income", y = "Density") +
  theme_minimal()
ggplotly(continous)


```

# Employment{.tabset}

## charts{.tabset}

### Graphs

```{r}


generate_sector_plot <- function(sector) {
  sector_sym <- sym(sector) # Convert string to symbol
  ggplot(merged_data, aes(x = MPO, y = !!sector_sym, fill = MPO)) +
    geom_bar(stat = "identity") +
    theme_minimal() +
    labs(title = paste("Employment in Sector", sector, "by MPO"), x = "MPO", y = "Number of Employees") +
    coord_flip()
}


long_data <- merged_data %>%
  filter(!is.na(MPO)) %>%
  pivot_longer(
    cols = starts_with("EMP_"), 
    names_to = "sector", 
    values_to = "employees"
  )

# Stacked bar plot for each sector with MPOs stacked
stacked_sectors_plot <- ggplot(long_data, aes(x = sector, y = employees, fill = MPO)) +
  geom_bar(stat = "identity", position = "stack") +  # Use position = "stack" for stacking
  theme_minimal() +
  labs(title = "Stacked Employment by Sector and MPO", x = "Employment Sector", y = "Number of Employees")
ggplotly(stacked_sectors_plot)


```

## Row

### Total Employment vs MPO

```{r}


# Total Employment by MPO
t=ggplot(merged_data, aes(x = MPO, y = TOTEMP, fill = MPO)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Total Employment by MPO", x = "MPO", y = "Total Employment") +
  coord_flip()

ggplotly((t))


```

### Total Employment vs Region

```{r}

# Convert REGION to a factor
merged_data$REGION <- as.factor(merged_data$REGION)

# Plot Total Employment by REGION
t <- ggplot(merged_data, aes(x = REGION, y = TOTEMP, fill = REGION)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Total Employment by REGION", x = "MPO", y = "Total Employment") +
  coord_flip()

# Convert ggplot object to an interactive plotly object

ggplotly(t)

```


### Household Employment in Tennessee

```{r}
library(ggplot2)
library(plotly)

# Assuming 'merged_data' is prepared and contains 'TOTEMP'

# Create the ggplot object
ggplot_map <- ggplot(data = merged_data) +
  geom_sf(aes(color = TOTEMP), size = 2.5) +  # Use color to represent TOTEMP
  scale_color_distiller(palette = "Spectral", direction = 1) +  # Adjust palette as needed
  labs(title = "Total Employment",
       subtitle = "By Traffic Analysis Zone or MPO",
       color = "Total Employment") +
  theme_minimal() +
  theme(legend.position = "bottom")

# Convert the ggplot object to plotly
interactive_map <- ggplotly(ggplot_map)

# Display the interactive map
interactive_map



```




# Land Usage{.tabset}

## charts{.tabset}

### Graphs

```{r}


attribute_data <- st_drop_geometry(merged_data)
land_use_data <- attribute_data %>%
  summarise(
    Residential = sum(Res, na.rm = TRUE),
    Commercial = sum(Com, na.rm = TRUE),
    Industrial = sum(Ind, na.rm = TRUE),
    Agricultural = sum(Agr, na.rm = TRUE),
    Vacant = sum(Vac, na.rm = TRUE)
  )%>%
  pivot_longer(
    cols = everything(),
    names_to = "Category",
    values_to = "Total_Land"
  )


# Now, create the bar plot
a <- ggplot(land_use_data, aes(x = Category, y = Total_Land, fill = Category)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Tennessee State Total Land Use by Category", x = "Category", y = "Total Land (sq units)")

ggplotly(a)




```

### land distribution by Region

```{r}


attribute_data <- st_drop_geometry(merged_data)
aggregated_data <- attribute_data %>%
  group_by(REGION) %>%
  summarise(
    Residential = sum(Res, na.rm = TRUE),
    Commercial = sum(Com, na.rm = TRUE),
    Industrial = sum(Ind, na.rm = TRUE),
    Agricultural = sum(Agr, na.rm = TRUE),
    Vacant = sum(Vac, na.rm = TRUE)
  )

# Reshape data for plotting with ggplot2
long_data <- aggregated_data %>% 
  pivot_longer(cols = c("Residential", "Commercial", "Industrial","Agricultural","Vacant"), names_to = "Land_Type", values_to = "Value")

# Create the bar plot
BAR<-ggplot(long_data, aes(x = REGION, y = Value, fill = Land_Type)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  theme_minimal() + 
  labs(title = "Land Types by Region", x = "REGION", y = "Total Land (sq units)") + 
  scale_fill_brewer(palette = "Set2") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggplotly(BAR)

```

### Landuse distribution vs MPO

```{r}


# Assuming 'merged_data' is your original dataframe

# Drop geometry column
attribute_data <- st_drop_geometry(merged_data)

# Group by MPO and calculate sum of each land type
aggregated_data <- attribute_data %>%
  filter(!is.na(MPO)) %>%  # Remove rows with null MPO
  group_by(MPO) %>%
  summarise(
    Residential = sum(Res, na.rm = TRUE),
    Commercial = sum(Com, na.rm = TRUE),
    Industrial = sum(Ind, na.rm = TRUE),
    Agricultural = sum(Agr, na.rm = TRUE),
    Vacant = sum(Vac, na.rm = TRUE)
  )

# Reshape data for plotting with ggplot2
long_data <- aggregated_data %>% 
  pivot_longer(cols = c("Residential", "Commercial", "Industrial","Agricultural","Vacant"), names_to = "Land_Type", values_to = "Value")

# Plotting
b <- ggplot(long_data, aes(x = MPO, y = Value, fill = Land_Type)) +
  geom_bar(stat = "identity", position = "stack") +
  theme_minimal() +
  labs(title = "Land Use Distribution by MPO", x = "MPO", y = "Area (sq units) ") +
  coord_flip()

# Convert ggplot to plotly object
plotly_plot <- ggplotly(b)

# Display the plotly plot
plotly_plot


```



## Row


### Residental land in Tennessee

```{r}

library(ggplot2)
library(plotly)

# Assuming 'merged_data' is prepared and contains 'TOTEMP'

# Create the ggplot object
ggplot_map <- ggplot(data = merged_data) +
  geom_sf(aes(color = Res), size = 2.5) +  # Use color to represent TOTEMP
  scale_color_distiller(palette = "Spectral", direction = 1) +  # Adjust palette as needed
  labs(title = "Residential Area",
       subtitle = "By Traffic Analysis Zone or MPO",
       ) +
  theme_minimal() +
  theme(legend.position = "bottom")

# Convert the ggplot object to plotly
interactive_map <- ggplotly(ggplot_map)

# Display the interactive map
interactive_map


```

### Commercial land in Tennessee

```{r}

library(ggplot2)
library(plotly)

# Assuming 'merged_data' is prepared and contains 'TOTEMP'

# Create the ggplot object
ggplot_map <- ggplot(data = merged_data) +
  geom_sf(aes(color = Com), size = 2.5) +  # Use color to represent TOTEMP
  scale_color_distiller(palette = "Spectral", direction = 1) +  # Adjust palette as needed
  labs(title = "Commercial Land",
       subtitle = "By Traffic Analysis Zone or MPO",
       ) +
  theme_minimal() +
  theme(legend.position = "bottom")

# Convert the ggplot object to plotly
interactive_map <- ggplotly(ggplot_map)

# Display the interactive map
interactive_map

```

### industrial land in Tennessee

```{r}

library(ggplot2)
library(plotly)

# Assuming 'merged_data' is prepared and contains 'TOTEMP'

# Create the ggplot object
ggplot_map <- ggplot(data = merged_data) +
  geom_sf(aes(color = Ind), size = 2.5) +  # Use color to represent TOTEMP
  scale_color_distiller(palette = "Spectral", direction = 1) +  # Adjust palette as needed
  labs(title = "Industrial Land",
       subtitle = "By Traffic Analysis Zone or MPO",
       ) +
  theme_minimal() +
  theme(legend.position = "bottom")

# Convert the ggplot object to plotly
interactive_map <- ggplotly(ggplot_map)

# Display the interactive map
interactive_map

```

### Agricultural land in Tennessee

```{r}

library(ggplot2)
library(plotly)

# Assuming 'merged_data' is prepared and contains 'TOTEMP'

# Create the ggplot object
ggplot_map <- ggplot(data = merged_data) +
  geom_sf(aes(color = Agr), size = 2.5) +  # Use color to represent TOTEMP
  scale_color_distiller(palette = "Spectral", direction = 1) +  # Adjust palette as needed
  labs(title = "Agricultural Land",
       subtitle = "By Traffic Analysis Zone or MPO",
       ) +
  theme_minimal() +
  theme(legend.position = "bottom")

# Convert the ggplot object to plotly
interactive_map <- ggplotly(ggplot_map)

# Display the interactive map
interactive_map

```



